---
title: "07_generate_baseline_data_2024"
author: "Sarah Willington"
date: "08-04-2024"
output: html_document
---

## info
markdown file to generate all the baseline data (no change in Linf)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load libraries and data, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)

project_dir <- here::here()
raw_data <- file.path(paste(project_dir, "0_data", "Prince_2023_MK_precise.csv", sep = "/"))
mk <- read_csv(raw_data) #read in raw data

```


## alfonsino baseline

```{r alfonsino - Beryx splendens }

mk_alfonsino <- mk %>% filter(species == "Beryx splendens")
summary(mk_alfonsino)

```

median m/k is 1.3429\
when m/k = 1.3429:\
Linf = 485.2\
k = 0.220\
t0 = -0.095\
m = 0.2954353\
max.age = 15.0\
LWa and LWb are NA\


```{r alfonsino baseline scenarios}
# Create the baseline data
baseline <- data.frame(
  species = "alfonsino",
  Linf = 485.2,
  K = 0.22,
  M = 0.2954353,
  t0 = -0.095,
  tmax = 15,
  scenario = "baseline"
)

# Create vectors of decreasing and increasing values for K and M (20% change steps)
decreasing_values <- c(0.8, 0.6, 0.4)
increasing_values <- c(1.2, 1.4, 1.6)

# Function to modify data
modify_data <- function(baseline, K_factor, M_factor) {
  return(baseline %>%
           mutate(
             K = K * K_factor,
             M = M * M_factor,
             scenario = "experimental"
           ))
} #close modify function

# Create an empty data frame to store the results
result_data <- data.frame()

# nested loops to loop through each combination of changing M and K ####

  # Loop through each value of K and M
  for (K_factor in decreasing_values) {
    for (M_factor in increasing_values) {
      modified_data <- modify_data(baseline, K_factor, M_factor)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M in reverse order
  for (K_factor in increasing_values) {
    for (M_factor in decreasing_values) {
      modified_data <- modify_data(baseline, K_factor, M_factor)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M increasing
  for (K_factor1 in increasing_values) {
    for (M_factor1 in increasing_values) {
      modified_data <- modify_data(baseline,  K_factor1, M_factor1)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M decreasing
  for (K_factor2 in decreasing_values) {
    for (M_factor2 in decreasing_values) {
      modified_data <- modify_data(baseline,  K_factor2, M_factor2)
      result_data <- rbind(result_data, modified_data)
    }
  } #end loop

# Reset row names in the result_data
row.names(result_data) <- NULL

# Bind with baseline
result_data <- rbind(result_data, baseline)

```


```{r alfonsino baseline write csv}
result_data <- result_data %>%  
  mutate(M = round(M, 3))

project_dir <- here::here()
al_baseline_path <- file.path(paste(project_dir, "03_results", "output1", "01_baseline_scenarios_alfonsino.csv", sep = "/"))
write_csv(result_data, al_baseline_path)
```


```{r alfonsino baseline increasing K scenarios}

rm(list=ls()) # clear workspace

# Create the baseline data
baseline <- data.frame(
  species = "alfonsino",
  Linf = 485.2,
  K = 0.22,
  M = 0.2954353,
  t0 = -0.095,
  tmax = 15,
  scenario = "baseline"
)

# vector of increasing K
increasing_values <- seq(1.01, 2, by=0.05)

# Function to modify data
modify_data <- function(baseline, K_factor) {
  return(baseline %>%
           mutate(
             K = K * K_factor,
             scenario = "experimental"
           ))
} #close modify function

# Create an empty data frame to store the results
result_data <- data.frame()

# nested loops to loop through each decrease in Linf and each combination of changing K ####

  # Loop through each value of K 
for (K_factor in increasing_values) {
      modified_data <- modify_data(baseline, K_factor)
      result_data <- rbind(result_data, modified_data)
 }#end loop

# Reset row names in the result_data
row.names(result_data) <- NULL

# Bind with baseline
result_data <- rbind(result_data, baseline)

```


```{r alfonsino baseline increasing K write csv}
result_data <- result_data %>%  
  mutate(M = round(M, 3))

project_dir <- here::here()
al_baselinek_path <- file.path(paste(project_dir, "03_results", "output1", "02_baselinek_scenarios_alfonsino.csv", sep = "/"))
write_csv(result_data, al_baselinek_path)
```



## swordfish baseline

```{r swordfish - Xiphias gladius }
rm(list=ls()) # clear workspace

project_dir <- here::here()
raw_data <- file.path(paste(project_dir, "0_data", "Prince_2023_MK_precise.csv", sep = "/"))
mk <- read_csv(raw_data) #read in raw data

#get swordfish data
mk_swordfish <- mk %>% filter(species == "Xiphias gladius")
summary(mk_swordfish)

```

median m/k is 1.0873\
There is no data for when M/k = 1.0873, but the closest (in direction of the mean M/k) is m/k = 1.1251:\
Linf = 1831.6\
k = 0.391\
t0 = -0.0210\
m = 0.4399305\
max.age = 10\

```{r swordfish baseline scenarios}
# Create the baseline data
baseline <- data.frame(
  species = "swordfish",
  Linf = 1831.6,
  K = 0.391,
  M = 0.4399305,
  t0 = -0.0210,
  tmax = 10,
  scenario = "baseline"
)

# Create vectors of decreasing and increasing values for K and M (20% change steps)
decreasing_values <- c(0.8, 0.6, 0.4)
increasing_values <- c(1.2, 1.4, 1.6)

# Function to modify data
modify_data <- function(baseline, K_factor, M_factor) {
  return(baseline %>%
           mutate(
             K = K * K_factor,
             M = M * M_factor,
             scenario = "experimental"
           ))
} #close modify function

# Create an empty data frame to store the results
result_data <- data.frame()

# nested loops to loop through each combination of changing M and K ####

  # Loop through each value of K and M
  for (K_factor in decreasing_values) {
    for (M_factor in increasing_values) {
      modified_data <- modify_data(baseline, K_factor, M_factor)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M in reverse order
  for (K_factor in increasing_values) {
    for (M_factor in decreasing_values) {
      modified_data <- modify_data(baseline, K_factor, M_factor)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M increasing
  for (K_factor1 in increasing_values) {
    for (M_factor1 in increasing_values) {
      modified_data <- modify_data(baseline,  K_factor1, M_factor1)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M decreasing
  for (K_factor2 in decreasing_values) {
    for (M_factor2 in decreasing_values) {
      modified_data <- modify_data(baseline,  K_factor2, M_factor2)
      result_data <- rbind(result_data, modified_data)
    }
  } #end loop

# Reset row names in the result_data
row.names(result_data) <- NULL

# Bind with baseline
result_data <- rbind(result_data, baseline)

```


```{r swordfish baseline write csv}
result_data <- result_data %>%  
  mutate(M = round(M, 3))

project_dir <- here::here()
sw_baseline_path <- file.path(paste(project_dir, "03_results", "output1", "01_baseline_scenarios_swordfish.csv", sep = "/"))
write_csv(result_data, sw_baseline_path)
```

```{r swordfish baseline increasing K scenarios}

rm(list=ls()) # clear workspace

# Create the baseline data
baseline <- data.frame(
  species = "swordfish",
  Linf = 1831.6,
  K = 0.391,
  M = 0.4399305,
  t0 = -0.0210,
  tmax = 10,
  scenario = "baseline"
)

# vector of increasing K
increasing_values <- seq(1.01, 2, by=0.05)

# Function to modify data
modify_data <- function(baseline, K_factor) {
  return(baseline %>%
           mutate(
             K = K * K_factor,
             scenario = "experimental"
           ))
} #close modify function

# Create an empty data frame to store the results
result_data <- data.frame()

# nested loops to loop through each decrease in Linf and each combination of changing K ####

  # Loop through each value of K 
for (K_factor in increasing_values) {
      modified_data <- modify_data(baseline, K_factor)
      result_data <- rbind(result_data, modified_data)
 }#end loop

# Reset row names in the result_data
row.names(result_data) <- NULL

# Bind with baseline
result_data <- rbind(result_data, baseline)

```


```{r swordfish baseline increasing K write csv}
result_data <- result_data %>%  
  mutate(M = round(M, 3))

project_dir <- here::here()
sw_baselinek_path <- file.path(paste(project_dir, "03_results", "output1", "02_baselinek_scenarios_swordfish.csv", sep = "/"))
write_csv(result_data, sw_baselinek_path)
```

## moon wrasse baseline

```{r moonwrasse - Thalassoma lunare }

rm(list=ls()) # clear workspace

project_dir <- here::here()
raw_data <- file.path(paste(project_dir, "0_data", "Prince_2023_MK_precise.csv", sep = "/"))
mk <- read_csv(raw_data) #read in raw data

mk_moonwrasse <- mk %>% filter(species == "Thalassoma lunare")
summary(mk_moonwrasse)

```

median m/k is 1.3391\
There is no data for when M/k = 1.3391, but the closest (in direction of the mean M/k) is m/k = 1.3453825:\
Linf = 141\
k = 0.540\
t0 = -0.136\
m = 0.7265065\
max.age = 6\

```{r moonwrasse baseline scenarios}
# Create the baseline data
baseline <- data.frame(
  species = "moonwrasse",
  Linf = 141,
  K = 0.540,
  M = 0.7265065,
  t0 = -0.136,
  tmax = 6,
  scenario = "baseline"
)

# Create vectors of decreasing and increasing values for K and M (20% change steps)
decreasing_values <- c(0.8, 0.6, 0.4)
increasing_values <- c(1.2, 1.4, 1.6)

# Function to modify data
modify_data <- function(baseline, K_factor, M_factor) {
  return(baseline %>%
           mutate(
             K = K * K_factor,
             M = M * M_factor,
             scenario = "experimental"
           ))
} #close modify function

# Create an empty data frame to store the results
result_data <- data.frame()

# nested loops to loop through each combination of changing M and K ####

  # Loop through each value of K and M
  for (K_factor in decreasing_values) {
    for (M_factor in increasing_values) {
      modified_data <- modify_data(baseline, K_factor, M_factor)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M in reverse order
  for (K_factor in increasing_values) {
    for (M_factor in decreasing_values) {
      modified_data <- modify_data(baseline, K_factor, M_factor)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M increasing
  for (K_factor1 in increasing_values) {
    for (M_factor1 in increasing_values) {
      modified_data <- modify_data(baseline,  K_factor1, M_factor1)
      result_data <- rbind(result_data, modified_data)
    }
  }
  # Loop through each value of K and M decreasing
  for (K_factor2 in decreasing_values) {
    for (M_factor2 in decreasing_values) {
      modified_data <- modify_data(baseline,  K_factor2, M_factor2)
      result_data <- rbind(result_data, modified_data)
    }
  } #end loop

# Reset row names in the result_data
row.names(result_data) <- NULL

# Bind with baseline
result_data <- rbind(result_data, baseline)

```


```{r moonwrasse baseline write csv}
result_data <- result_data %>%  
  mutate(M = round(M, 3))

project_dir <- here::here()
mw_baseline_path <- file.path(paste(project_dir, "03_results", "output1", "01_baseline_scenarios_moonwrasse.csv", sep = "/"))
write_csv(result_data, mw_baseline_path)
```

```{r moonwrasse baseline increasing K scenarios}

rm(list=ls()) # clear workspace

# Create the baseline data
baseline <- data.frame(
  species = "moonwrasse",
  Linf = 141,
  K = 0.540,
  M = 0.7265065,
  t0 = -0.136,
  tmax = 6,
  scenario = "baseline"
)

# vector of increasing K
increasing_values <- seq(1.01, 2, by=0.05)

# Function to modify data
modify_data <- function(baseline, K_factor) {
  return(baseline %>%
           mutate(
             K = K * K_factor,
             scenario = "experimental"
           ))
} #close modify function

# Create an empty data frame to store the results
result_data <- data.frame()

# nested loops to loop through each decrease in Linf and each combination of changing K ####

  # Loop through each value of K 
for (K_factor in increasing_values) {
      modified_data <- modify_data(baseline, K_factor)
      result_data <- rbind(result_data, modified_data)
 }#end loop

# Reset row names in the result_data
row.names(result_data) <- NULL

# Bind with baseline
result_data <- rbind(result_data, baseline)

```


```{r moonwrasse baseline increasing K write csv}
result_data <- result_data %>%  
  mutate(M = round(M, 3))

project_dir <- here::here()
mw_baselinek_path <- file.path(paste(project_dir, "03_results", "output1", "02_baselinek_scenarios_moonwrasse.csv", sep = "/"))
write_csv(result_data, mw_baselinek_path)
```



